Tweets

### 2020-05-29 Rmd/jse-special-issue-2020.Rmd
switched to using memorial day tweets

In line with recent calls for students to work with real data [@nolan2010computing], our 
collection of tweets has served as a valuable resource in our mentoring 
of undergraduate data science research. 
Working with real data allows students to develop proficiency not 
only in statistical analysis,
but also in related data science skills, including data 
transfer from online sources, data storage, using data 
from multiple file formats, and communicating findings and their limitations. 
Collaboratively asking and addressing novel questions with 
our collection of tweets gave mentored students 
opportunities to develop competency in all of these areas.

@@

In our experience, student interests vary, and their initial ability to articulate research goals may be limited. An initial brainstorming session may clarify their interests and encourage them to think critically about goals under the time constraints of their academic schedules. Additionally, we anticipate that sharing completed student project reports will guide student thinking about the scope of possible projects.

We mentored one student whose interest in financial time series guided her project. A second student formulated a project around event detection from tweet time series.

@@

Latent Dirichlet allocation is a statistical method for inferring latent (unobservable) topics (or themes)
from a corpus (or collection) of documents. We pretend that
there's an imaginary process for creating documents in the corpus. For each document,
we choose a discrete distribution over topics. For example, some Mother's Day tweets wish 
mothers a happy celebration. This may constitute one topic in the corpus. 
Having chosen a distribution over topics, we then select document words by first 
drawing a topic from the distribution over topics, then drawing a word from the
chosen topic. 
Thus, a topic is technically defined as a distribution over words in a fixed vocabulary (or collection of words).

The literature on latent Dirichlet allocation and related methods is vast, 
and we won't attempt to review it here.

@@

We sought to validate our hypothesis that we could detect a major social 
media event by examining tweet topic content at distinct time periods. 
As a proof of principle of our event detection strategy, we chose to 
analyze tweets during and after Mother's Day 2020. We fitted latent Dirichlet allocation models for each of four 
distinct five-minute periods. The first period began at noon Eastern time on Mother's Day 2020. Subsequent time periods started 
24, 48, and 72 hours later. 
We defined each time period to be a single collection, or corpus, of tweets. We then fitted latent 
Dirichlet allocation models to each corpus.

We used several criteria to evaluate latent Dirichlet allocation model fits, with 
emphasis on choosing a reasonable number of topics per corpus.
Our strategy involved both visualization and more quantitative approaches to model 
evaluation. For every model, we created one word cloud per topic.

@@

## Appendix 2: LDA results for May 12, 13, and 14

```{r, fig.cap = "Top terms for LDA model from May 12, 2020"}
knitr::include_graphics("../results/beta-2020-05-12.png")
```

```{r, fig.cap = "Top terms for LDA model from May 13, 2020"}
knitr::include_graphics("../results/beta-2020-05-13.png")
```

```{r, fig.cap = "Top terms for LDA model from May 14, 2020"}
knitr::include_graphics("../results/beta-2020-05-14.png")
```

@@

## Appendix 3: R analysis code

>>>>

In line with recent calls for students to work with real data [@nolan2010computing;@carver2016guidelines], our 
collection of tweets has served as a valuable resource in our mentoring 
of undergraduate data science research. 
Working with real data allows students to develop proficiency not 
only in statistical analysis,
but also in related data science skills, including data 
transfer from online sources, data storage, using data 
from multiple file formats, and communicating findings. 
Collaboratively asking and addressing novel questions with 
our collection of tweets gave mentored students 
opportunities to develop competency in all of these areas.

@@

In our experience, student interests vary, and students' ability to articulate research goals 
may be limited. An initial brainstorming session may clarify their interests and encourage 
them to think critically about goals under the time constraints of their academic schedules. 
Additionally, we anticipate that sharing completed student project reports will guide student 
thinking about the scope of possible projects.

We mentored one student whose interest in financial time series and tweet sentiment analysis 
guided her project. A second student formulated a project around event detection from tweet 
time series.

@@

Latent Dirichlet allocation is a statistical method for inferring latent (unobservable) topics (or themes)
from a corpus (or collection) of documents. We pretend that
there's an imaginary process for creating documents in the corpus. For each document,
we choose a discrete distribution over topics. For example, some tweets from Memorial Day may refer to the holiday. This may constitute one topic in the corpus. 
Having chosen a distribution over topics, we then select document words by first 
drawing a topic from the distribution over topics, then drawing a word from the
chosen topic. 
Thus, a topic is technically defined as a distribution over words in a fixed vocabulary (or collection of words) [@blei2003latent].

@@

We sought to validate our hypothesis that we could detect a major social 
media event by examining tweet topic content at distinct time periods. 
As a proof of principle of our event detection strategy, we chose to 
analyze tweets before, during, and after Memorial Day 2020 (May 25, 2020). We fitted latent Dirichlet allocation models for each of three
distinct five-minute periods. The first period began at noon Eastern time on May 24, 2020. Subsequent time periods started 
24 and 48 hours later. 
We defined each time period to be a single collection, or corpus, of tweets. We then fitted latent 
Dirichlet allocation models to each corpus.

One challenge in latent Dirichlet allocation modeling is the choice of the 
number of topics per corpus. To address this, we fitted each corpus to models

We used several criteria to evaluate latent Dirichlet allocation model fits, with 
emphasis on choosing a small number of topics per corpus.
Our strategy involved both visualization and more quantitative approaches to model 
evaluation. For every model, we created one word cloud per topic.

@@

@@

## Appendix 2: R analysis code

### 2020-05-29 Rmd/tweets-one.Rmd
switched to using memorial day tweets

```{r, recover-{{date}}}
fn <- paste0("../data/", {{date}}, "-analyzed-tweet-ids.csv")
tr_fn <- paste0("../data/", {{date}}, "-rehydrated-tweets.csv")
if (!file.exists(fn)){

  tweets <- recover_stream(path = "{{tweet_json_file}}", verbose = TRUE) 
  tweets %>%
    dplyr::select(status_id) %>%
    readr::write_csv(path = paste0("../data/tweets-status-ids-", "{{date}}", ".csv"), 
                     col_names = FALSE)
  broken_ids <- readr::read_csv("broken_tweets.txt", col_names = FALSE) %>%
    dplyr::rename(status_id = X1) %>%
    dplyr::mutate(status_id = trimws(format(status_id, scientific = FALSE)))
  ids <- readr::read_csv(paste0("../data/tweets-status-ids-", "{{date}}", ".csv"), col_names = FALSE) %>%
    dplyr::rename(status_id = X1) %>%
    dplyr::mutate(status_id = as.character(status_id))
  tweets_rehydrated1 <- rtweet::lookup_tweets(statuses = ids$status_id, 
                                              parse = TRUE)
  tweets_rehydrated2 <- rtweet::lookup_tweets(statuses = broken_ids$status_id, parse = TRUE)
  tweets_rehydrated <- tweets_rehydrated1 %>%
    dplyr::bind_rows(tweets_rehydrated2)
  # save ids of tweets that we actually used
  readr::write_csv(tibble::as_tibble(tweets_rehydrated$status_id), path = fn, col_names = FALSE)
  readr::write_csv(tweets_rehydrated, path = tr_fn)
}
```

@@

```{r, lda-{{date}}}
tweets_lda <- topicmodels::LDA(tweets_dtm, k = 8)
tweets_beta <- tidytext::tidy(tweets_lda, matrix = "beta")
```

```{r, topterms-{{date}}}
tweets_top_terms <- tweets_beta %>%
  dplyr::group_by(topic) %>%
  dplyr::top_n(10, beta) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(topic, -beta)
```

```{r, plot-{{date}}}
tweets_top_terms %>%
  dplyr::mutate(term = tidytext::reorder_within(term, beta, topic)) %>%
  ggplot2::ggplot(ggplot2::aes(term, beta, fill = factor(topic))) +
  ggplot2::geom_col(show.legend = FALSE) +
  ggplot2::facet_wrap(~ topic, scales = "free") +
  ggplot2::coord_flip() + 
  theme(axis.text.x = element_text(angle = 270))

@@

```{r, ggsave-{{date}}}
ggplot2::ggsave(paste0("../results/beta-", {{date}}, ".png"))
```

```{r, sleep-{{date}}}
Sys.sleep(15 * 60) # sleep for 15 minutes to allow api limits to reset
```

>>>>

```{r, recover-{{date}}}
fn <- paste0("../data/", "{{date}}", "-analyzed-tweet-ids.csv")
#fn <- paste0("../data/", "2020-05-10", "-analyzed-tweet-ids.csv")
tr_fn <- paste0("../data/", "{{date}}", "-rehydrated-tweets.rds")
#tr_fn <- paste0("../data/", "2020-05-10", "-rehydrated-tweets.rds")
sleep <- FALSE
if (!file.exists(tr_fn)){
  sleep <- TRUE
  tweets <- recover_stream(path = "{{tweet_json_file}}", verbose = TRUE) 
  #tweets <- recover_stream(path = "../data/2020-05-26 12:00:02-tweets.json", verbose = TRUE) 
  id_fn <- paste0("../data/tweets-status-ids-", "{{date}}", ".csv")
  #id_fn <- paste0("../data/tweets-status-ids-", "2020-05-26", ".csv")
  tweets %>%
    dplyr::select(status_id) %>%
    readr::write_csv(path = id_fn, 
                     col_names = FALSE)
  broken_ids <- readr::read_csv("broken_tweets.txt", col_names = FALSE) %>%
    dplyr::rename(status_id = X1) %>%
    dplyr::mutate(status_id = trimws(format(status_id, scientific = FALSE)))
  ids <- readr::read_csv(id_fn, col_names = FALSE) %>%
    dplyr::rename(status_id = X1) %>%
    dplyr::mutate(status_id = as.character(status_id))
  tweets_rehydrated1 <- rtweet::lookup_tweets(statuses = ids$status_id, 
                                              parse = TRUE)
  tweets_rehydrated2 <- rtweet::lookup_tweets(statuses = broken_ids$status_id, parse = TRUE)
  tweets_rehydrated <- tweets_rehydrated1 %>%
    dplyr::bind_rows(tweets_rehydrated2)
  # save ids of tweets that we actually used
  readr::write_csv(tibble::as_tibble(tweets_rehydrated$status_id), path = fn, col_names = FALSE)
  saveRDS(tweets_rehydrated, tr_fn)
} else {
  tweets_rehydrated <- readRDS(tr_fn)
}
```

@@

## Comparing models with distinct number of topics

```{r, tuning-{{date}}}
(result <- ldatuning::FindTopicsNumber(
  tweets_dtm,
  topics = seq(from = 5, to = 40, by = 1),
  metrics = c("CaoJuan2009", "Arun2010", "Deveaud2014"),
  method = "VEM",
  mc.cores = 2L,
  verbose = TRUE
))
```

```{r, tuning2-{{date}}}
ldatuning::FindTopicsNumber_plot(result)
```

```{r, tuning-fig-{{date}}}
png(paste0("../results/ldatuning-", "{{date}}", ".png"))
ldatuning::FindTopicsNumber_plot(result)
dev.off()
```

```{r, optimal-k-{{date}}}
(optimal_k_CaoJuan2009 <- result$topics[which.min(result$CaoJuan2009)])
(optimal_k_Arun2010 <- result$topics[which.min(result$Arun2010)])
(optimal_k_Deveaud2014 <- result$topics[which.max(result$Deveaud2014)])
```

```{r, lda-CJ09-{{date}}}
tweets_lda <- topicmodels::LDA(tweets_dtm, k = optimal_k_CaoJuan2009)
tweets_beta <- tidytext::tidy(tweets_lda, matrix = "beta")
```

```{r, topterms-CJ09-{{date}}}
tweets_top_terms <- tweets_beta %>%
  dplyr::group_by(topic) %>%
  dplyr::top_n(10, beta) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(topic, -beta)
```

```{r, plot-CJ09-{{date}}}
tweets_top_terms %>%
  dplyr::mutate(term = tidytext::reorder_within(term, beta, topic)) %>%
  ggplot2::ggplot(ggplot2::aes(term, beta, fill = factor(topic))) +
  ggplot2::geom_col(show.legend = FALSE) +
  ggplot2::facet_wrap(~ topic, scales = "free") +
  ggplot2::coord_flip() + 
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 270))

@@

```{r, ggsave-CJ09-{{date}}}
ggplot2::ggsave(paste0("../results/beta-CJ09", "{{date}}", ".png"))
```

### Arun2010

```{r, lda-Arun2010-{{date}}}
tweets_lda <- topicmodels::LDA(tweets_dtm, k = optimal_k_Arun2010)
tweets_beta <- tidytext::tidy(tweets_lda, matrix = "beta")
```

```{r, topterms-Arun2010-{{date}}}
tweets_top_terms <- tweets_beta %>%
  dplyr::group_by(topic) %>%
  dplyr::top_n(10, beta) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(topic, -beta)
```

```{r, plot-Arun2010-{{date}}}
tweets_top_terms %>%
  dplyr::mutate(term = tidytext::reorder_within(term, beta, topic)) %>%
  ggplot2::ggplot(ggplot2::aes(term, beta, fill = factor(topic))) +
  ggplot2::geom_col(show.legend = FALSE) +
  ggplot2::facet_wrap(~ topic, scales = "free") +
  ggplot2::coord_flip() + 
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 270))

```{r, ggsave-Arun2010-{{date}}}
ggplot2::ggsave(paste0("../results/beta-Arun2010", "{{date}}", ".png"))
```

### Deveaud2014

```{r, lda-Deveaud2014-{{date}}}
tweets_lda <- topicmodels::LDA(tweets_dtm, k = optimal_k_Deveaud2014)
tweets_beta <- tidytext::tidy(tweets_lda, matrix = "beta")
```

```{r, topterms-Deveaud2014-{{date}}}
tweets_top_terms <- tweets_beta %>%
  dplyr::group_by(topic) %>%
  dplyr::top_n(10, beta) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(topic, -beta)
```

```{r, plot-Deveaud2014-{{date}}}
tweets_top_terms %>%
  dplyr::mutate(term = tidytext::reorder_within(term, beta, topic)) %>%
  ggplot2::ggplot(ggplot2::aes(term, beta, fill = factor(topic))) +
  ggplot2::geom_col(show.legend = FALSE) +
  ggplot2::facet_wrap(~ topic, scales = "free") +
  ggplot2::coord_flip() + 
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 270))

```{r, ggsave-Deveaud2014-{{date}}}
ggplot2::ggsave(paste0("../results/beta-Deveaud2014", "{{date}}", ".png"))
```

### Wait

```{r, sleep-{{date}}}
if (sleep) Sys.sleep(15 * 60) # sleep for 15 minutes to allow api limits to reset
```

### 2020-05-29 Rmd/tweets.Rmd
Removed 'goals' section
switched to using memorial day tweets

## Goals

1. Write a R script to query twitter streaming api to get 1 percent sample over 5'.
1. Demonstrate how to set up a crontab to execute script every 5'.
1. Cautions about twitter databases: deletions and amendments to tweets

@@

```{r}
date <- c("2020-05-10", "2020-05-11", "2020-05-12", "2020-05-13", "2020-05-14")
tweet_json_file <- c("../data/2020-05-10 12:00:01-tweets.json",
  "../data/2020-05-11 12:00:01-tweets.json", 
  "../data/2020-05-12 12:00:02-tweets.json",
  "../data/2020-05-13 12:00:02-tweets.json", 
  "../data/2020-05-14 12:00:02-tweets.json")
out <- purrr::pmap_dfc(tibble::tibble(date, tweet_json_file),
                       .f = function(date, tweet_json_file){
                         knitr::knit_expand(text = readLines("tweets-one.Rmd"))
                       })
```

## May 10, 2020

@@

## May 11, 2020

@@

## May 12, 2020

@@

## May 13, 2020

`r knitr::knit(text = out[[4]])`

## May 14, 2020

`r knitr::knit(text = out[[5]])`

>>>>

@@

```{r}
date <- c("2020-05-24", "2020-05-25", "2020-05-26")
tweet_json_file <- c("../data/2020-05-24 12:00:02-tweets.json",
  "../data/2020-05-25 12:00:02-tweets.json", 
  "../data/2020-05-26 12:00:02-tweets.json"
  )
out <- purrr::pmap_dfc(tibble::tibble(date, tweet_json_file),
                       .f = function(date, tweet_json_file){
                         knitr::knit_expand(text = readLines("tweets-one.Rmd"))
                       })
```

## May 24, 2020

@@

## May 25, 2020

@@

## May 26, 2020

@@
